---
title: "Approach to process user upload data"
author: "Kristina Riemer"
date: '2023-01-20'
output: html_document
---

## Purpose

Users will upload raw data, which will have length and/or weight measurements with one row per observation (a single fish or multiple fish summed). These data need to be summarized to the three metrics of interest for each unique combination of area, species, collection method, type of water body, and year. The three metrics are: 

1. [Catch per unit effort](https://en.wikipedia.org/wiki/Catch_per_unit_effort) (CPUE)
2. Length frequency
3. Relative weight

## Defining user inputs

1. CPUE requires `effort` column
2. Length frequency requires `total_length` column 
3. Relative weight requires `weight` and `total_length` columns

Required columns in input dataframe: 

- **
- **Area**: need at least one of these columns, can have all three
  - `type` is "all" or "state"
  - `area` can only be "North_America"
  - `state` must be spelled out and capitalized
  - `ecoregion` comes from [EPA level 1 ecoregions](https://www.epa.gov/eco-research/ecoregions-north-america)
- **Date**: only a numeric `year` column is needed
- **Measurements**: 
  - `total_length` is fish record length (mm)
  - `weight` is fish record weight (g)
  - `effort` is specified in **Collection method**
- **Collection method**: see the table below for details
  - `method` must exactly match one of the options in the first column
  - `effort` will contain a numeric value that has correct type and unit

| **Method name**          | **Effort type** | **Unit** |
|--------------------------|-----------------|----------|
| boat_electrofishing      | Time            | seconds  |
| tow_barge_electrofishing | Time            | seconds  |
| raft_electrofishing      | Time            | seconds  |
| trawl                    | Time            | seconds  |
| gill_net_fall            | Number of nets  | number   |
| gill_net_fall            | Number of nets  | number   |
| hoop_net                 | Number of nets  | number   |
| small_catfish_hoopnet    | Number of nets  | number   |
| large_catfish_hoopnet    | Number of nets  | number   |
| seine                    | Number of nets  | number   |
| bag_seine                | Number of nets  | number   |
| stream_seine             | Number of nets  | number   |
| backpack_electrofishing  | Area            | m2       |
| snorkel                  | Area            | m2       |

- **Type of water body**
  - `waterbody_type` must be one of the following: "large_standing_waters", "small_standing_waters", "two_story_standing_waters", "wadeable_streams", "rivers"
- **Species** must be in `common_name` as species common name, which must match one of those listed below, as from `FSA::PSDlit`:

```{r, echo=FALSE}
species_list <- FSA::PSDlit
print(species_list$species)
```

## Prepare to calculate metrics

### Read in libraries
```{r, warning=FALSE, message=FALSE}
library(readxl)
library(dplyr)
library(stringr)
library(FSA)
library(ggplot2)
```

### Read in and modify example datasets

Example of final dataset: 

```{r, echo=FALSE}
process_example <- read.csv("input_examples/AFS_processed_test_data.csv")
head(process_example)
```

Read in example dataset. 

```{r}
raw_y1 <- readxl::read_xlsx("input_examples/AFS_test_data.xlsx") %>%
  tidyr::separate(date, into = c("month", "day", "year"), sep = c(1, 3, 7)) %>% 
  mutate(year = as.numeric(year), 
         common_name = str_to_title(common_name), 
         waterbody_type = str_replace_all(waterbody_type, " ", "_"))
```

Simulate extra data for additional years for raw data. 

```{r}
raw_y2 <- raw_y1 %>% 
  mutate(year = year + 1, 
         total_length = total_length + 50, 
         effort = effort / 2)
```

Simulate North America data. 

```{r}
raw_y3 <- raw_y2 %>% 
  mutate(state = "North America", 
         common_name = "Bluegill")
```

Simplify dataset to include only necessary columns. 

```{r}
raw_example <- bind_rows(raw_y1, raw_y2, raw_y3) %>% 
  rename(area = state) %>% 
  select(area, waterbody_name, common_name, method, waterbody_type, year, effort, effort_transect, transect, total_length, weight) %>% 
  mutate(type = case_when(area == "Arizona" ~ "state", 
                          area == "North America" ~ "all"))
```

Save example user input dataset. 

```{r}
write.csv(raw_example, "input_examples/uu_raw.csv", row.names = FALSE)
```


## Functions to calculate metrics

### 1. CPUE

For time-based methods, convert effort from seconds to hours. Calculate CPUE by dividing number of fish records for each species by effort. 

```{r}
# average each lake, then calculate metric (number of fish and effort), then do average by state

# get sampleID first = state, waterbody name, year, method
# then average measurements by sampleID
  # measurements = effort
# n is for number of sampleIDs in a state/ecoregion

non_net_methods <- c("boat_electrofishing", "tow_barge_electrofishing", 
                     "raft_electrofishing", "small_mesh_trawl", "large_mesh_trawl")
net_methods <- c("gill_net_spring", "stream_seine", "hoop_net", 
                 "small_catfish_hoopnet", "drifting_trammel_net",
                 "large_catfish_hoopnet", "seine", "bag_seine")
bp_methods <- c("backpack_electrofishing")

calculate_cpue <- function(df){
  stopifnot(c("type", "area", "common_name", "method", "waterbody_type", 
              "year", "effort") %in% colnames(df)) #todo check these
  cpue <- df %>% 
    group_by(type, area, waterbody_name, year, method, waterbody_type, common_name) %>% 
    summarize(n_fish = n(), 
              effort_sampleID = mean(effort)) %>% 
    mutate(effort = case_when(method %in% non_net_methods ~ effort_sampleID / 3600, 
                              method %in% net_methods ~ effort_sampleID)) %>% 
    mutate(cpue = n_fish / effort) %>% # n is number of fish
    group_by(type, area, common_name, method, waterbody_type) %>%
    summarize_at(vars(cpue), list(mean = mean, se = se,
                                `5%` = ~quantile(x = ., probs = 0.05),
                                `25%` = ~quantile(x = ., probs = 0.25),
                                `50%` = ~quantile(x = ., probs = 0.50),
                                `75%` = ~quantile(x = ., probs = 0.75),
                                `95%` = ~quantile(x = ., probs = 0.95)),
               na.rm = TRUE) %>%
    mutate(metric = "CPUE")
  return(cpue)
}

cpue_example <- calculate_cpue(raw_example)

```

### 2. Length frequency

Calculate Gabelhouse category for each fish record using length and species identity. For each species and each year, determine proportion of records in each category. Then calculate mean and standard error of length frequency for each species by category across years. 

```{r}
# lf by lake then averaged across state
calculate_lf <- function(df){
    stopifnot(c("type", "area", "common_name", "method", "waterbody_type", 
                "year", "total_length") %in% colnames(df))
  lf <- raw_example %>% 
    mutate(gcat = psdAdd(total_length, common_name, what = "incremental")) %>% 
    group_by(type, area, common_name, method, waterbody_type, year) %>% 
    count(gcat) %>% 
    mutate(lenfreq = n / sum(n)) %>% 
    group_by(type, area, common_name, method, waterbody_type, gcat) %>% 
    summarize_at(vars(lenfreq), list(mean = mean, se = se), na.rm = TRUE) %>% 
    mutate(metric = "Length Frequency", 
           gcat = factor(gcat, levels = c("stock", "quality", "preferred",
                                          "memorable", "trophy")))
  return(lf)
}
```

### 3. Relative weight

Calculate Gabelhouse category for each fish record using length and species identity, and relative weight for each fish record using width, length, and species identity. Then calculate mean and standard error for relative weight for each species by category across years. 

```{r}
calculate_rw <- function(df){
  stopifnot(c("type", "area", "common_name", "method", "waterbody_type", "year", 
              "total_length", "weight") %in% colnames(df))
  rw <- raw_example %>% 
    mutate(gcat = psdAdd(total_length, common_name, what = "incremental"), 
         relweight = wrAdd(wt = weight, len = total_length, spec = common_name)) %>% 
    group_by(type, area, common_name, method, waterbody_type, gcat) %>% 
    summarize_at(vars(relweight), list(mean = mean, se = se, 
                                `5%` = ~quantile(x = ., probs = 0.05),
                                `25%` = ~quantile(x = ., probs = 0.25),
                                `50%` = ~quantile(x = ., probs = 0.50),
                                `75%` = ~quantile(x = ., probs = 0.75),
                                `95%` = ~quantile(x = ., probs = 0.95)), na.rm = TRUE) %>% 
    mutate(metric = "Relative Weight", 
           gcat = factor(gcat, levels = c("stock", "quality", "preferred",
                                          "memorable", "trophy")))
  return(rw)
}
```

## Generate metrics dataset

Run 3 functions on user input data, then combine into final dataset. 

TODO: add count

```{r}
counts <- raw_example %>% 
  group_by(type, area, common_name, method, waterbody_type) %>%
  summarise(N = n())
```


```{r}
cpue_example <- calculate_cpue(raw_example)
lf_example <- calculate_lf(raw_example)
rw_example <- calculate_rw(raw_example)

final_example <- bind_rows(cpue_example, lf_example, rw_example) %>% 
  left_join(counts, by = c("type", "area", "common_name", "method", "waterbody_type"))
```

### Plot metrics dataset

```{r}
ggplot(cpue_example, aes(y = area)) +
        geom_boxplot(aes(xmin = `5%`,
                         xlower = `25%`,
                         xmiddle = `50%`,
                         xupper = `75%`,
                         xmax = `95%`),
                     stat = "identity") +
  facet_wrap(vars(common_name))
```

```{r}
ggplot(lf_example, aes(x = gcat, y = mean)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(drop = FALSE) +
  facet_wrap(vars(common_name)) +
        geom_errorbar(aes(ymin = mean - se,
                          ymax = mean + se),
                      width = 0)
```

```{r}
ggplot(rw_example, aes(x = gcat)) +
  geom_point(aes(y = mean)) +
  geom_line(aes(y = mean)) +
  scale_x_discrete(drop = FALSE) +
  facet_wrap(vars(common_name))
```
